---
title: "Bayes Theorem"
subtitle: "Applied To Data Analysis"
author: "Andy Grogan-Kaylor"
institute: "University of Michigan"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{css, echo=FALSE}

@import url('https://fonts.googleapis.com/css2?family=Montserrat&display=swap');
.title-slide { 
  color: #ffcb05;
  background-color: #00274C; 
}
.title-slide h1 {
    color: #ffcb05;
}
pre {
  white-space: pre-wrap;
}
h1, h2, h3 {
  font-family: 'Montserrat', sans-serif;
}
body { 
    font-family: 'Montserrat', sans-serif;
}
.author, .date {
  font-family: 'Montserrat', sans-serif;
}
blockquote { 
  border-left: 5px solid #ffcb05; 
  margin: 1.5em 10px; 
  padding: 0.5em 10px;
}

```

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

xaringanExtra::use_tile_view()

xaringanExtra::use_animate_css()

xaringanExtra::use_share_again()

xaringanExtra::use_progress_bar(color = "#ffcb05", location = "bottom")

```

# How To Navigate This Presentation

* Use the <span style="font-size:100px">&#8678;</span> and <span style="font-size:100px">&#8680;</span> keys to move through the presentation.
* Press *o* for *panel overview*.

---
class: animated, fadeIn
# Derivation

--

|               | D=1                  | D=0                             |
|:--------------|:---------------------|:--------------------------------|
| H=1           | $P(D, H)$            | $P(\text{not} D, H)$            |
| H=0           | $P(D, \text{not} H)$ | $P(\text{not} D, \text{not} H)$ |

--

From the definition of conditional probability: 

--

$P(D|H) = P(D,H) / P(H)$

--

$P(H|D) = P(D,H) / P(D)$

---
class: animated, fadeIn
# Then: 

--

$P(D|H)P(H) = P(D,H)$

--

$P(H|D)P(D) = P(D,H)$

--

# Then:

$P(D|H)P(H) = P(H|D)P(D)$

---
class: animated, fadeIn
# Bayes Theorem:

$P(H|D) = \frac{P(D|H) P(H)}{P(D)}$ 

--
# In Words:

$\text{posterior} \propto \text{likelihood} \times \text{prior}$ 

```{r, echo=FALSE, eval=FALSE}

# Expanded Versions

$P(H|D) = \frac{P(D|H)P(H)}{P(D|H)P(D) + P(D|\neg H)P(\neg D)}$ 

--
  
$P(D|H) = \frac{P(H|D)P(D)}{\Sigma P(H|D_i)P(D_i)}$
  
--

$P(D|H) = \frac{P(H|D)P(D)}{\int P(H|D_i)P(D_i) dD}$


```

---
class: animated, fadeIn
# Example 

Consider an example using 1,000 hypothetical studies. We imagine that only 10% of interventions are likely to have results. We adopt standard assumptions of adopting an $\alpha$, or chance of detecting an effect when one is not there of 5%. We similarly assume 80% power $\beta$, or a 20% chance of failing to detect an effect when it is not present. 

--

|  Data (D)                | D=1 (effect)        | D=0 (no effect)    |
|:-------------------------|:--------------------|:-------------------|
| Hypothesis (H)           | 100 effects         | 900 non-effects    |
| H=1 (conclude effect)    | 80 true positives   | 45 false positives |
| H=0 (conclude no effect) | 20 false negatives  | 855 true negatives |

--

> With thanks to the [Wikipedia article](https://en.wikipedia.org/wiki/Bayes%27_theorem#Cancer_rate) on this topic for inspiration for this example.

---
class: animated, fadeIn, center, top
# Visualization

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.retina=3}

library(ggplot2)

library(ggmosaic)

library(waffle)

library(hrbrthemes)

library(michigancolors)

library(tidyr)

library(patchwork)

D <- c("yes", "no", "yes", "no") # data

D <- factor(D, levels = c("yes", "no")) # order of levels

H <- c("yes", "yes", "no", "no") # hypothesis

# H <- factor(H, levels = c("yes", "no")) # order of levels

mycount <- c(80, 45, 20, 855) # count

mylabel <- c("true positive", 
             "false positive", 
             "false negative", 
             "true negative")

mylabel <- factor(mylabel, 
                  levels = c("true positive", 
                             "false positive", 
                             "false negative", 
                             "true negative"))

Bayesdata <- data.frame(D, H, mycount, mylabel)

p1 <- Bayesdata %>%
  ggplot(aes(fill = mylabel, values = mycount)) + 
  geom_waffle(n_rows = 25, 
              color = "grey",
              make_proportional = FALSE,
              height = 0.8, 
              width = 0.8) +
  # facet_wrap(~mylabel) +
  labs(title = "1,000 Hypothetical Studies") +
  scale_fill_manual(name = "Category",
                    values = c(michigancolors("blue"),
                               michigancolors("maize"),
                               michigancolors("tappan red"),
                               michigancolors("wave field green"))) +
  coord_fixed() +
  theme(plot.title = element_text(size=rel(3)),
        legend.text = element_text(size = rel(1)),
        aspect.ratio = .5) +
  # theme_ipsum_rc(grid="") +
  theme_minimal() +
  theme_enhance_waffle() 

p1 

```


---
class: animated, fadeIn
# Calculations

--

$$P(\text{H=1} | \text{D=1}) = \frac{P(\text{D=1} | \text{H=1}) P(\text{H=1})}{P(\text{D})}$$

--

$$=\frac{P(\text{D=1} | \text{H=1}) P(\text{H=1})}{P(\text{D=1} | \text{H=1}) P(\text{H=1}) + P(\text{D=0} | \text{H=1}) P(\text{H=1})}$$

--

$$P(\text{H=1} | \text{D=1}) = \frac{.8 \times .1}{.08 + .045} = .64$$

--

> See also [Thinking Through Bayesian Ideas](https://agrogan.shinyapps.io/Thinking-Through-Bayes/)

---
class: animated, fadeIn
# Discussion

--
* Calculations suggest that a true effect is likely in 64% of the cases where one concludes the presence of an effect.

--

* Consequently, calculations suggest that 36% of the time, when one concludes there is an effect, there is actually no effect. 

--

* Put another way, despite setting $\alpha = .05$, 36% of cases result in a false positive.

--

* Notice how the Bayesian approach lets us estimate the probability of different hypotheses given the data. In some cases, this may afford us the opportunity to accept our null hypothesis $H_0$.



