---
title: "The Value of Accepting the Null Hypothesis"
author: "Andy Grogan-Kaylor"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    highlight: haddock
    latex_engine: xelatex
  html_document: 
    css: ../styles.css
    highlight: haddock
institute: University of Michigan
bibliography: ../Bayes.bib
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

library(knitr)

```

# Background

In standard frequentist models, we cannot formally accept the Null Hypothesis $H_0$, but can only reject, or fail to reject, $H_0$.

Bayesian models allow one to both accept and reject $H_0$ [@Kruschke2018]. 

Accepting $H_0$ may have consequences for affirming similarity, universality, or treatment invariance [@Morey2018]. 

# Important Substantive Cases

The Value of Accepting the Null Hypothesis $H_0$

```{r}

library(pander) # nice tables

case <- c("Equivalence Testing",
          "Equivalence Testing",
          "Retiring Interventions",
          "Contextual Equivalence",
          "Family Member Equivalence",
          "Full Mediation",
          "Theory Simplification")

description <- c("Equivalence Of 2 Treatments Or Interventions",
                 "Equivalence Of 2 Groups On An Outcome",
                 "There Is No Evidence That Intervention X Is Effective",
                 "Equivalence of a Predictor Across Contexts (Moderation)",
                 "Equivalence of a Predictor Across Family Members",
                 "$x \\rightarrow y$ Association Is Completely Mediated; No Direct Effect",
                 "Removing An Association From A Theory")

H0 <- c("$\\beta_1 = \\beta_2$",
        "$\\bar{x_1} = \\bar{x_2}$",
        "$\\beta_{intervention} = 0$",
        "$\\beta_{interaction} = 0$",
        "$\\beta_{parent1} = \\beta_{parent2}$",
        "$\\beta_{xmy} \\neq 0$; $\\beta_{xy} = 0$",
        "$\\beta_x = 0$")

example <- c("The effect of Treatment 1 is indistinguishable from the effect of Treatment 2 (especially important if one treatment is much more expensive, or time consuming than another).",
             "Men and women are more similar than different *wrt* psychological processes [@Hyde2005].", 
             "Evidence consistently suggests that a particular treatment has near zero effect.",
             "Warm and supportive parenting is equally beneficial across different contexts or countries.",
             "Parenting from one parent is equivalent to parenting from another parent",
             "The relationship of the treatment and the outcome is completely mediated by mechanism *m*.",
             "There is no evidence that x is associated with y.")

Bayesdata <- data.frame(case, 
                     description,
                     H0,
                     example)

save(Bayesdata, file = "accepting-H0.RData")

pander(Bayesdata, split.cells = 10)

```

# References






